---
title: "Ectopic Expression"
output: html_document
date: "2025-01-10"
---

## Setup
```{r setup, include=FALSE}

library(knitr)
library(rmarkdown)
library(tidyverse)
library(readxl)
library(magrittr)
library(DescTools)
library(ggthemes)
library(ggpubr)
library(rcartocolor)
library(Cairo)
library(gtools)
library(car)
library(rstatix)
knitr::opts_chunk$set(echo = TRUE)
source("../Utils/theme_Publication.R")
source("../Utils/wormpal.R")
```

# AFD-rGC Ecoptic Expression Analysis
Analysis and plotting for a series of experiments testing the effect of 
misepxresion of *Ss-gcy-23.1*, *Ss-gcy-23.2*, and *Ss-gcy-23.3* in *C. elegans* ASE neurons, in response to the pseudo-fictive PTe 20-40$^\circ$C ramp.   

## Data Inputs
Data are preprocessed using Matlab TCI file `TCI_analyzeEctopic.m`, which calculates the temperature at which point an experimental trace rises above 3*rms of mean control trace for at least 40 seconds. 

This calculation is run independently for each trace, and the value (in $^\circ$C)
is saved.

```{r message=FALSE}
dat_ectopic_file <- c('../Data/Yi_EctopicExpression_R_Import.xlsx')
if (is_empty(dat_ectopic_file)){
  dat_ectopic_file <- file.choose()
}
meta_ectopic <-read_excel(dat_ectopic_file, sheet = "Metadata")
dat_ectopic <- read_excel(dat_ectopic_file, sheet = 1) %>%
  group_by(rGC,Strain)

dat_ectopic_filtered <- dat_ectopic %>%
  dplyr::filter(rGC != "ASE_ctrl") %>%
  dplyr::filter(!is.na(Thresh_temp)) %>%
  dplyr::bind_rows(dplyr::filter(dat_ectopic, rGC == "ASE_ctrl"))
```

## rGC Plot Templates 

```{r}

plot_ectopic_noctrl <- function(dat, y, meta, ylim, xlabel,
                                ylabel, stat.test, 
                                stat.y.position,
                                title = NULL){
  
  ggplot(data = dat, mapping = aes(x = rGC, y = y)) + 
    
    stat_boxplot(geom = "errorbar", 
                 width = 0.25,
                 linewidth = 0.5) +
    geom_boxplot(width = 0.5,
                 outlier.shape = NA,
                 fatten = 1,
                 key_glyph = "rect",
                 alpha = 0.3,
                 aes(fill = rGC)) +
    geom_jitter(size = 2,
                stroke = .75,
                width = .2,
                aes(color = rGC, shape = Strain)) +
    stat_pvalue_manual(stat.test, 
                       y.position = stat.y.position,
                       step.increase = 0.1,
                       tip.length = 0,
                       hide.ns = TRUE,
                       bracket.shorten = 0.2,
                       label = "p.adj.signif"
    ) + 
    scale_y_continuous(limits = ylim, 
                       expand = expansion(add = 0)) +
    xlab(xlabel) +
    ylab(ylabel) +
    ggtitle(title) +
    scale_fill_manual(values = worm.pal()) +
    scale_color_manual(values = worm.pal())+
    scale_shape_manual(values = rep(c(21:23), 20))  +
    theme_Publication() +
    theme(aspect.ratio = 4/3,
          axis.text.x = element_text(angle= 45,
                                     
                                     hjust = 1))
}
```


## Temperature Threshold  
Stats are Kruskal-Wallis Test with Dunn's Post tests for difference in response amplitude.  
``` {r}
# ---- Temperature Threshold Plot ----

xlabel <- "Ss-AFD-rGCs"
ylabel <- " Temperature (C)"
title <- "Threshold Temperature"
ylim <- c(19, 42)

dat.subset <-dplyr::filter(dat_ectopic_filtered, rGC != "ASE_ctrl")%>%
  dplyr::mutate(Strain = as_factor(Strain)) %>%
  dplyr::mutate(rGC = as_factor(rGC))

# Kruskal-Wallis testing of response amplitude differences
one.way <- kruskal.test(Thresh_temp ~ rGC, 
                        data = dat.subset)

# Posthoc Dunn tests with Bonferroni adjustment
post.hoc <- dunn.test::dunn.test(dat.subset$Thresh_temp, 
                                 dat.subset$rGC, 
                                 method = "bonferroni",
                                 table = FALSE,
                                 list = TRUE)

stat.test.ectopic.thresh <- dplyr::tibble(
  comparisons = post.hoc$comparisons,
  p.adj = post.hoc$P.adjusted,
  p.adj.signif = stars.pval(post.hoc$P.adjusted)
) %>%
  mutate(p.adj.signif = case_when(p.adj.signif == " " ~ "ns",
                                  TRUE ~ as.character(p.adj.signif))) %>%
  tidyr::separate(comparisons, 
                  into = c("group1", "group2"),
                  sep = " - ")

p_ectopic_thresh<- plot_ectopic_noctrl(dat.subset, 
                                       dat.subset$Thresh_temp, 
                                       meta_ectopic,
                                       ylim,
                                       xlabel, 
                                       ylabel,
                                       stat.test.ectopic.thresh,
                                       42,
                                       title)


print(p_ectopic_thresh)


```

## Temperature driving maximum response  

Stats are Kruskal-Wallis Test with Dunn's Post tests for difference in response amplitude.  
```{r}
# Kruskal-Wallis testing of response amplitude differences
one.way <- kruskal.test(Tmax ~ rGC, 
                        data = dat.subset)

# Posthoc Dunn tests with Bonferroni adjustment
post.hoc <- dunn.test::dunn.test(dat.subset$Tmax, 
                                 dat.subset$rGC, 
                                 method = "bonferroni",
                                 table = FALSE,
                                 list = TRUE)

xlabel <- "Ss-AFD-rGCs"
ylabel <- "Temperature (deg C)"
title <- "Temperature driving max calcium response"
ylim <- c(19,42)

stat.test.ectopic.Tmax <- dplyr::tibble(
  comparisons = post.hoc$comparisons,
  p.adj = post.hoc$P.adjusted,
  p.adj.signif = stars.pval(post.hoc$P.adjusted)
) %>%
  mutate(p.adj.signif = case_when(p.adj.signif == " " ~ "ns",
                                  TRUE ~ as.character(p.adj.signif))) %>%
  tidyr::separate(comparisons, 
                  into = c("group1", "group2"),
                  sep = " - ")

p_Tmax<- plot_ectopic_noctrl(dat.subset, 
                             dat.subset$Tmax, 
                             meta_ectopic,
                             ylim,
                             xlabel, 
                             ylabel,
                             stat.test.ectopic.Tmax,
                             42,
                             title)

print(p_Tmax)
```

## Group and Save Plots
```{r}
ggarrange(
  p_ectopic_thresh,
  p_Tmax, 
  ncol = 2,
  nrow = 1
) %>%
  ggsave(filename ="../Outputs/YiEctopicrGCExpression.pdf",
         device = cairo_pdf,
         width = 9,
         # height = 3.5,
         path = "./")
```
